---
title: "Finger millet blast structure output - pohelper"
author: "Joseph"
date: "14/10/2019"
output: html_document
---


## Installing software

Requires R version 3.5 or > # You can use conda to get your R to the latest version
Uncomment to install software
```{r setup, include=TRUE, echo=FALSE}
##
# install.packages(c("devtools","ggplot2","gridExtra","gtable","label.switching","tidyr"),dependencies=T)

## **install pophelper package from CRAN**
# install.packages('pophelper')

## **install pophelper package from GitHub**
# devtools::install_github('royfrancis/pophelper')

## check version
packageDescription("pophelper", fields="Version")

```

## Loading the libraries 


```{r libraries, include=TRUE ,echo=TRUE}
library("pophelper")
library("ggplot2")
library("gridExtra")
library("gtable")
library("label.switching")
library("tidyr")
library("devtools")

```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Setting up working directory

```{r results_directory, echo=FALSE}

#setwd("/Users/ATEMIA/Desktop/Results_all_runs/Results")

sfiles <- list.files(path=("C:/Users/icipe/Downloads/Ubuntu-Share/WindowsSubLinux/Home/fmb-research/structure/structure-2020/results/"), full.names = T )

slist <- readQ(files=sfiles)

```


## Getting delta K

Atleast two runs are required for dermination of delter K
```{r evannoMethod, echo=TRUE}
sr1 <- summariseQ(tabulateQ(slist))
p <- evannoMethodStructure(data=sr1,exportplot=T,returnplot=T,returndata=F,basesize=20,width=300,height=300,units="mm",linesize=1.5,
                           writetable=F)
grid.arrange(p)

```

## Plotting

#### Colors
```{r, echo=F}
clist <- list(
"standard_12"=c("#2121D9","#9999FF","#DF0101","#04B404","#FFFB23","#FF9326","#A945FF","#0089B2","#B26314","#610B5E","#FE2E9A","#BFF217"),
"rich.colors"=pophelper:::getColours(13),"shiny"=c("#1D72F5","#DF0101","#77CE61","#FF9326","#A945FF","#0089B2","#FDF060","#FFA6B2","#BFF217","#60D5FD","#CC1577",
                                                   "#F2B950","#7FB21D","#EC496F","#326397","#B26314","#027368","#A4A4A4","#610B5E"),
"strong"=c("#11A4C8","#63C2C5","#1D4F9F","#0C516D","#2A2771","#396D35","#80C342","#725DA8","#B62025","#ED2224","#ED1943","#ED3995","#7E277C","#F7EC16","#F8941E",
           "#8C2A1C","#808080"),
"oceanfive"=c("#00A0B0", "#6A4A3C", "#CC333F", "#EB6841", "#EDC951"),
"keeled"=c("#48B098", "#91CB62", "#FFEE3B", "#FB9013", "#FF3C28"),
"vintage"=c("#400F13", "#027368", "#A3BF3F", "#F2B950", "#D93A2B"),
"muted"=c("#46BDDD","#82DDCE","#F5F06A","#F5CC6A","#F57E6A"),
"teal"=c("#CFF09E","#A8DBA8","#79BD9A","#3B8686","#0B486B"),
"merry"=c("#5BC0EB","#FDE74C","#9BC53D","#E55934","#FA7921"),
"funky"=c("#A6CEE3", "#3F8EAA", "#79C360", "#E52829", "#FDB762","#ED8F47","#9471B4"),
"retro"=c("#01948E","#A9C4E2","#E23560","#01A7B3","#FDA963","#323665","#EC687D"),
"cb_paired"=c("#A6CEE3","#1F78B4","#B2DF8A","#33A02C","#FB9A99","#E31A1C","#FDBF6F","#FF7F00","#CAB2D6","#6A3D9A","#FFFF99","#B15928"),
"cb_set3"=c("#8DD3C7","#FFFFB3","#BEBADA","#FB8072","#80B1D3","#FDB462","#B3DE69","#FCCDE5","#D9D9D9","#BC80BD","#CCEBC5","#FFED6F"),
"morris"=c("#4D94CC","#34648A","#8B658A","#9ACD32","#CC95CC","#9ACD32","#8B3A39","#CD6601","#CC5C5B","#8A4500"),
"wong"=c("#000000","#E69F00","#56B4E9","#009E73","#F0E442","#006699","#D55E00","#CC79A7"),
"krzywinski"=c("#006E82","#8214A0","#005AC8","#00A0FA","#FA78FA","#14D2DC","#AA0A3C","#FA7850","#0AB45A","#F0F032","#A0FA82","#FAE6BE"))

```

### Align K

Reordering the clusters suitably

When multiple repeats are run for each K, the order of clusters may be jumbled for each run.
Therefore, when plotting multiple runs within each K.

```{r, echo=TRUE}

#p1 <- plotQ(alignK(slist[c(1:22)]),imgoutput="join",returnplot=T,exportplot=F,quiet=T,basesize=11)
p1 <- plotQ(alignK(sortQ(slist)[c(41:45)]),imgoutput = "join",returnplot=T,exportplot=F,
            quiet=T,basesize=16,clustercol=clist$shiny,showsp = F,
            showlegend=T,showindlab=T,useindlab=T,showyaxis=F,panelspacer=0.3,sharedindlab=T,indlabangle=-90,indlabsize=2,
            indlabspacer=-1,showtitle=T,showsubtitle=T,titlelab="STRUCTURE PLOT",width=300,height=300,units="mm",
            subtitlelab="K9 Plots", legendkeysize=10,legendtextsize=10,legendmargin=c(2,2,2,0),legendrow=1)

grid.arrange(p1$plot[[1]])

```

- alignK() first aligns all runs within each K.
- Then progressively aligns pairs of K from lowest to the highest. 
- Cluster align within K is performed using the stephens algorithm implemented in the R package label.switching.
- Aligning across K is carried out using the idea implemented in CLUMPAK. 
- CLUMPP software can also be used for aligning clusters.


```{r, echo=TRUE}

#p1 <- plotQ(alignK(slist[c(1:22)]),imgoutput="join",returnplot=T,exportplot=F,quiet=T,basesize=11)
p1 <- plotQ(mergeQ(alignK(sortQ(slist))[c(41:45)]),returnplot=T,exportplot=T,quiet=T,basesize=16,clustercol=clist$shiny,
            showlegend=T,showindlab=T,useindlab=T,showyaxis=F,panelspacer=0.3,sharedindlab=T,indlabangle=-90,indlabsize=3,
            indlabspacer=-1,showtitle=T,showsubtitle=T,titlelab="STRUCTURE PLOT",titlehjust = 0.5,width=1000,height=100,
            units="mm",splab = "K=9",subtitlelab="Merged OLD K9 Plots",legendkeysize=10,legendtextsize=10,
            legendmargin=c(2,2,2,0),legendrow=1)

grid.arrange(p1$plot[[1]])


```


### plotQ()



###
```{r, echo=T}
p1 <- plotQ(alignK(sortQ(slist[c(41,44)])),imgoutput="join",returnplot=T,exportplot=F,quiet=T,basesize=11,
            showlegend=T,showindlab=T,useindlab=T,showyaxis=F,panelspacer=0.3,sharedindlab=T,indlabangle=-90,indlabsize=2,indlabspacer=-1,
            legendkeysize=10,legendtextsize=10,legendmargin=c(2,2,2,0),legendrow=1,clustercol=clist$krzywinski,
            splab=paste0("K=",sapply(slist[c(46,49)],ncol)),
            showtitle=T,showsubtitle=T,titlelab="STRUCTURE PLOT",
            subtitlelab="K9 Plots")

grid.arrange(p1$plot[[1]])

```


### Sorting
```{r, echo=TRUE} 
p1 <- plotQ(alignK(sortQ(slist[c(41:45)])),imgoutput = "join",returnplot=T,exportplot=T,quiet=T,basesize=11, showlegend=T,showindlab=T,
            useindlab=T,sortind="all",showyaxis=F,panelspacer=1,sharedindlab=F,indlabangle=-90,indlabsize=5,legendkeysize=10,legendtextsize=10,
            legendmargin=c(2,2,2,0),legendrow=1,width=400,height=400, clustercol=clist$shiny,splab=paste0("K=",sapply(slist[c(46:50)],ncol)),
            showtitle=T,showsubtitle=T,titlelab="STRUCTURE K9 RUNS ",titlecol="#800020",titlehjust = 0.5,
            subtitlelab="Sorted by cluster",subtitlecol="#800020")

grid.arrange(p1$plot[[1]])

```

### Multiline barplot
```{r, echo=TRUE}
p <- plotQMultiline(slist[4],exportplot=F,returnplot=T,spl=120,lpp=4)
grid.arrange(p$plot[[1]][[1]])

```

#### Labelling and sorting
```{r, echo=TRUE}
k<-1
for (i in seq(from=1, to=4)){
# get a qlist
list1 <- slist[i]

# read some individual labels
inds <- read.delim(("~/Desktop/blast_isolate_labs copy.txt"),
                   header=FALSE, stringsAsFactors=FALSE)

# add labels as rownames to qlist
rownames(list1[[1]]) <- inds$V1


plotQMultiline(alignK(list1),useindlab=T,showlegend=T,showindlab=T,showyaxis=T,
               indlabcol="darkgreen", sortind="all",indlabangle=45, indlabvjust=1, indlabhjust=1,
               showtitle = T, titlelab = "EASTERN AFRICA FINGER MILLET BLAST STRUCTRE PLOT",titlecol = "#800020 ",showsubtitle = T,
               subtitlelab = paste("K = 9 - run",k) ,titlehjust = 0.5, subtitlecol = "#800020",height=12, mar=c(0,0,0,0.5),clustercol=clist$shiny)
k<-k+1
}
```


#### 2 Labelling and sorting
```{r, echo=TRUE}
# uncomment the loop to print multiple runs

for (i in seq(from=28, to=30)){
  
# get a qlist
list1 <- slist[c(28,29,30)]

# read some individual labels
inds <- read.delim(("/Volumes/Untitled/OUT_15k_30K"),
                   header=FALSE, stringsAsFactors=FALSE)

# add labels as rownames to qlist
rownames(list1[[1]]) <- inds$V1


plotQMultiline(mergeQ(alignK(list1)),useindlab=T,showlegend=T,showindlab=T,showyaxis=T,
               indlabcol="darkgreen", sortind="all",indlabangle=45, indlabvjust=1, indlabhjust=1,
               showtitle = T, titlelab = "EASTERN AFRICA FINGER MILLET BLAST STRUCTRE PLOT",titlecol = "#800020",showsubtitle = T,
               subtitlelab = "Merged K = 9",titlehjust = 0.5, subtitlecol = "#800020",height=12, mar=c(0,0,0,0.5),clustercol=clist$shiny)
#}
```


### testing sorting
```{r, echo=TRUE}
# get a qlist
list1 <- slist[c(46:50)]

# read some individual labels
inds <- read.delim(("~/Desktop/blast_isolate_labs copy.txt"),
                   header=FALSE, stringsAsFactors=FALSE)

# add labels as rownames to qlist
rownames(list1[[1]]) <- inds$V1

p1 <- plotQ(alignK(sortQ(slist[c(46:50)])),imgoutput = "join",returnplot=T,exportplot=T,quiet=T,basesize=11, showlegend=T,showindlab=T,
            useindlab=T,sortind="all",showyaxis=F,panelspacer=1,sharedindlab=F,indlabangle=-90,indlabsize=5,legendkeysize=10,
            legendtextsize=10,legendmargin=c(2,2,2,0),legendrow=1,width=400,height=100,units="mm",
            clustercol=clist$shiny,splab=paste0("K=",sapply(slist[c(46:50)],ncol)),
            showtitle=T,showsubtitle=T,titlelab="STRUCTURE RUN 1 to 5",titlecol="#800020",titlehjust = 0.5,
            subtitlelab="Sorted by cluster k=4",subtitlecol="#800020")

grid.arrange(p1$plot[[1]])

```


### Exporting plots from different runs
```{r, echo=TRUE}

k=1 # naming the different runs
# depending on the k values you want printed out, the seq is specified
for (i in seq(from=46, to =50)){
# get a qlist
list1 <- slist[(i)]

# read some individual labels
inds <- read.delim(("/Volumes/Untitled/OUT_15k_30K"),
                   header=FALSE, stringsAsFactors=FALSE)

# add labels as rownames to qlist
rownames(list1[[1]]) <- inds$V1

plotQ(alignK((list1)),returnplot=T,exportplot=T,quiet=T,basesize=20, showlegend=T,showindlab=T,
            useindlab=T,showyaxis=T,panelspacer=0.5,sharedindlab=F,
            indlabangle=90,indlabcol= "darkgreen", indlabsize = 9,sortind = "all",
            legendkeysize=20,legendtextsize=20,legendmargin=c(4,4,4,0),legendrow=1,width=700,height=100,units="mm",
            clustercol=clist$shiny,splab=paste0("K=",sapply(list1,ncol)),
            showtitle=T,showsubtitle=T,titlelab=paste("STRUCTURE RUN",k) ,titlecol="#800020",titlehjust = 0.5,
            subtitlelab="Sorted by cluster",subtitlecol="#800020")
k <- k+1
#grid.arrange(p1$plot[[1]])
}
```